---
title: "Intro to ggplot2: A grammar for graphics"

output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
theme: lumen
---

### Introducation

`ggplot2` is a graphing library in R that makes beautiful graphs. It is a R package created by R programmer Hadley Wickham. ggplot2 is one of the most widely used R packages. Moreover, it builds data graphics incrementally from small pieces of code.


The first time you want to use a package, you need to install it. To do so, you type:
```{r  eval=F}
install.packages('ggplot2')
```

You then need to activate by executing the command:
```{r  eval=F}
library(ggplot2)
```

Each time you start an R session you will need to activate the package(s) you need, but you don’t need
to install them again. 


### Data Applied 

We will be using the data set CIACountries from the `mdsr` package. First, we have to `intall` and `library` the `mdsr` package.

```{r eval=F}
install.packages('mdsr')
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(mdsr)
```


The CIACountries data table contains the following seven variables for 236 countries:

+ population (pop)
+ area (area)
+ gross domestic product (gdp)
+ percentage of GDP spent on education (educ)
+ length of roadways per unit area (roadways)
+ internet use as a fraction of the population (net_users)
+ number of barrels of oil produced per day (oil_prod)

We can apply the `head` function to havea a look of the CIACountries data table. The `%>%knitr::kable()` command just make is part of table looks better than before.

```{r}
head(CIACountries) %>% knitr::kable()
```


The ToothGrowth data table contains the following seven variables for 60 guinea pigs:

+ Tooth length (len)
+ Supplement type (supp)
+ Dose in milligrams/day (dose)

```{r}
data("ToothGrowth")

# Load data and convert dose to a factor variable
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# add a outlier to the dataset
ToothGrowth[1,1] <- 60

head(ToothGrowth) %>% knitr::kable()
```

### 1.ggplot Syntax

Unlike base graphics, ggplot works with dataframes and not individual vectors. The `ggplot()`command creates a plot, with the argument data telling us which data to use. The `aes()` command within ggplot() indicates the desired x and
y variables.

```{r fig.height=3.5, fig.width=4}
ggplot(data = CIACountries, aes(y=gdp, x = educ))
```

Although datafram, x and y are specified, there are no points or lines in it. This is because I have only told ggplot what dataset to use and what columns should be used for X and Y axis. I haven’t explicitly asked it to draw any points.


>___The following are the Geoms( Bars, lines, text, points) options to apply on the plot___

+ `geom_bar()` : creates a layer with bars representing different statistical properties
+ `geom_point()` : creates a layer with data points (as on a scatterplot)
+ `geom_line()` : creates a layer with a straight line
+ `geom_smooth()` : creates a layer with a ‘smoother’ (a line that summarizes the data as a whole rather than connecting individual points)
+ `geom_histogram()` : creates a layer with a histogram
+ `geom_boxplot()` : creates a layer with a box-whisker diagram
+ `geom_text()` : creates a layer with text in it
+ `geom_errorbar()` : creates a layer with error bars in it
+ `geom_hline()` and `geom_vline()` : creates a layer with a user-defined horizontal or vertical line respectively.

<br>

### 2. Scatter plot

#### 2.1 Basic Scatter plot

Let’s make a scatterplot by adding points using a geom layer called `geom_point`. The `color`, `size` and `shape` can be applied to in the `geom_point` to change the color, size and shape of the points in scatter plot

```{r eval=F}
# Basic Scatter plot
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point()

# Change the point color
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point(color="steelblue")

# Change the point size, and shape
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point(size=2, shape=23)
```
```{r echo=FALSE, fig.height=3.5, fig.width=10,message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

p1 <- ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point()

# Change the point color
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point(color="steelblue")

# Change the point size, and shape
p3 <- ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point(size=2, shape=23)

grid.arrange(p1, p2, p3, ncol=3)
```

#### 2.2 Change the point color/size automatically

The command `size`, `shape` and `color` are used to set size, shape and color vary based on state categories. Users can decide to set one, two or all of them.

`guide_legend(title="")` is used to add the legend name. If size has been set to vary based on state categories, the command of add legend name is `size=guide_legend(title="")`. If color has been set to vary based on state categories, the command of add legend name is `color=guide_legend(title="")`.

```{r eval=F}
# Change the point size
ggplot(data = CIACountries, aes(y=gdp, x = educ))  + geom_point(aes(size = roadways)) +
  guides(size=guide_legend(title="Roadway"))

# Change the point size and coloe
ggplot(data = CIACountries, aes(y=gdp, x = educ))  + geom_point(aes(color = net_users, size = roadways)) +
  guides(color=guide_legend(title="Internet Use"), size=guide_legend(title='Roadway'))
```


```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Change the point size
p1 <- ggplot(data = CIACountries, aes(y=gdp, x = educ))  + geom_point(aes(size = roadways)) +
  guides(size=guide_legend(title="Roadway"))

# Change the point size and coloe
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ))  + geom_point(aes(color = net_users, size = roadways)) +
  guides(color=guide_legend(title="Internet Use"), size=guide_legend(title='Roadway'))

grid.arrange(p1, p2, ncol=2)
```

#### 2.3 Label points in the scatter plot

The function `geom_text()` can be used :

```{r fig.height=3.5, fig.width=5.5,message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = CIACountries, aes(y=gdp, x = educ))  + geom_text(aes(label = country), size = 3) 
```

#### 2.4.1 Add regression lines

The functions `geom_smooth()` can be used to add regression lines to a scatter plot 

```{r eval=F}
geom_smooth(method="auto", se=TRUE, fullrange=FALSE, level=0.95)
```

+ `method`: smoothing method to be used. Possible values are lm, glm, gam, loess, rlm. The default value is loess.
+ `se`: If TRUE, confidence interval is displayed around smooth.
+ `fullrange`: If TRUE, the fit spans the full range of the plot
+ `level`: level of confidence interval to use. Default value is 0.95

```{r eval=F}
# Add the regression line
ggplot(data = CIACountries, aes(y=gdp, x = educ))  +  geom_point()+
  geom_smooth(method=lm)

# Remove the confidence interval
ggplot(data = CIACountries, aes(y=gdp, x = educ))  +  geom_point()+
  geom_smooth(method=lm, se=FALSE)
```

```{r  echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)
# Add the regression line
p1 <- ggplot(data = CIACountries, aes(y=gdp, x = educ))  +  geom_point()+
  geom_smooth(method=lm)

# Remove the confidence interval
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ))  +  geom_point()+
  geom_smooth(method=lm, se=FALSE)

grid.arrange(p1, p2, ncol=2)
```

#### 2.4.2 Change the points and lines

`shape` and `color` in the ` geom_point()` can chance the color of shape of the points in scatter plot. `color` and `linetype` in the ` geom_smooth()` can change color and linetype of the line.

```{r fig.height=3.5, fig.width=5.5,message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = CIACountries, aes(y=gdp, x = educ))  +  geom_point(shape=10, color="steelblue")+
  geom_smooth(method=lm,se=FALSE, linetype="dashed", color="orange")
```

#### 2.4.3 Multiple regression lines

In the scatter plot, users can plot more than one regression line in the plot by the following code. User need to set color change automatically by the command `color`. Then user need to have the command `geom_smooth()` to add the regression lines. Then there would be multiple regression lines in one graph.

Command set `fullrange` equals TRUE can extend the regression lines in the graph.

```{r eval=F}
ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm)+
  guides(color=guide_legend(title="Internet Use"))

# Remove confidence intervals
# Extend the regression lines
ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))
```
```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)
p1 <- ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm)+
  guides(color=guide_legend(title="Internet Use"))

# Remove confidence intervals
# Extend the regression lines
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))

grid.arrange(p1, p2, ncol=2)
```

#### 2.5.1 Change the point color/size manually

In previous section, if there exist multiple regression lines in the graph, the color and points would change automatically. The following command can set the color, size and shape manually:

+ `scale_shape_manual()`: for changeing point shapes manually
+ `scale_color_manual()`: for changeing point colors manually
+ `scale_size_manual()`: for changeing point sizes manually

The following example shows how to set the point and line color and size manually.

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))+
  scale_color_manual(values=c('#999999','#E69F00', '#56B4E9','#0072B2', '#D55E00', '#CC79A7'))+
  scale_size_manual(values=c(1,2,3,4,5))
```


#### 2.5.2 Change the point color/size manually by color group

It is also possible to change manually point and line colors using the functions :

+ `scale_color_brewer()` : to use color palettes from RColorBrewer package
+ `scale_color_grey()` : to use grey color palettes

```{r eval=F}
# Use brewer color palettes
ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))+
  scale_color_brewer(palette="Dark2")

# Use grey scale
ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))+
  scale_color_grey()
```
```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Use brewer color palettes
p1 <- ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))+
  scale_color_brewer(palette="Dark2")

# Use grey scale
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ, color=net_users))  + 
  geom_point()+
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE)+
  guides(color=guide_legend(title="Internet Use"))+
  scale_color_grey()

grid.arrange(p1, p2, ncol=2)
```


### 3. Box plot

#### 3.1 Basic Box plot

Let’s make a box plot by adding command `geom_boxplot()`. Set `notch` equals TRUE can change the shape of the box plot. 

```{r eval=F}
# Basic box plot
ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()

# Notched box plot
ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot(notch=TRUE)
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Basic box plot
p1 <- ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()

# Notched box plot
p2 <- ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot(notch=TRUE)

grid.arrange(p1, p2, ncol=2)
```


#### 3.2 Change Outlier and add mean points to a box plot

`geom_boxplot()` allowed user to change color, shape and zie of the outlier points by the following commands.

+ `outlier.shape()`: for changeing outlier point shapes manually
+ `outlier.colour()`: for changeing outlier point colors manually
+ `outlier.size()`: for changeing outlier point sizes manually

The command `stat_summary()` can be used to add mean points to a box plot

```{r eval=F}
# Change outlier, color, shape and size
ggplot(data = ToothGrowth, aes(y=len, x = supp))) + geom_boxplot()+
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4)

# add mean points to a box plot
ggplot(data = ToothGrowth, aes(y=len, x = supp))) + geom_boxplot()+
  stat_summary(fun.y=mean, geom="point", shape=23, size=4)
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Change outlier, color, shape and size
p1 <- ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  geom_boxplot(outlier.colour="red", outlier.shape=8, outlier.size=4)

# add mean points to a box plot
p2 <- ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  stat_summary(fun.y=mean, geom="point", shape=23, size=4)

grid.arrange(p1, p2, ncol=2)
```


#### 3.3 Box plot with dots

The points can be added to box plot by comman `geom_dotplot()` and `geom_jitter()`.

```{r eval=F}
# Box plot with dot plot
ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.3)

# Box plot with jittered points
# 0.2 : degree of jitter in x direction
ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  geom_jitter(shape=16, position=position_jitter(0.2))
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Box plot with dot plot
p1 <-ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.3)

# Box plot with jittered points
# 0.2 : degree of jitter in x direction
p2 <- ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  geom_jitter(shape=16, position=position_jitter(0.2))

grid.arrange(p1, p2, ncol=2)
```


#### 3.4 Select item to display

The command `scale_x_discrete()` can use to select the box to display on the boxplot. 

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = ToothGrowth, aes(y=len, x = supp)) + geom_boxplot()+
  scale_x_discrete(limits=c("OJ"))
```


#### 3.5.1 Change box plot colors automatically

The command `color` is used to set the line color of box plot vary based on state categories.\n 
The command `fill` is used to set the box color of box plot vary based on state categories.

```{r eval=F}
# Change box plot line colors by groups
ggplot(data = ToothGrowth, aes(y=len, x = supp color=supp)) + geom_boxplot()

# Change box plot colors by groups
ggplot(data = ToothGrowth, aes(y=len, x = supp, fill=supp)) + geom_boxplot()
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Change box plot line colors by groups
p1 <- ggplot(data = ToothGrowth, aes(y=len, x = supp, color=supp)) + geom_boxplot()

# Change box plot colors by groups
p2 <- ggplot(data = ToothGrowth, aes(y=len, x = supp, fill=supp)) + geom_boxplot()

grid.arrange(p1, p2, ncol=2)
```

>Boxplot Side by side

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = ToothGrowth, aes(x = dose, y = len)) + geom_boxplot(aes(fill = supp))
```


#### 3.5.2 Change box plot colors manually

The command `color` is used to set the line color of box plot.\\
The command `fill` is used to set the box color of box plot.\\
The command `scale_fill_manual` is used to set the box color of of each box.

```{r eval=F}
# Use single color
ggplot(ToothGrowth, aes(x = dose, y = len)) + geom_boxplot(fill="steelblue", color="black")

# Use custom color palettes
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Use single color
p1 <- ggplot(ToothGrowth, aes(x = dose, y = len)) + geom_boxplot(fill="steelblue", color="black")

# Use custom color palettes
p2 <- ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))

grid.arrange(p1, p2, ncol=2)
```


__It is also possible to change manually point and line colors using the functions :__

+ `scale_color_brewer()` : to use color palettes from RColorBrewer package
+ `scale_color_grey()` : to use grey color palettes

```{r eval=F}
# use Blues color palettes
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  scale_fill_brewer(palette="Blues") 

# Use grey scale
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  scale_fill_grey() + theme_classic() 
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# use brewer color palettes
p1 <-ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  scale_fill_brewer(palette="Blues")  

# Use grey scale
p2 <- ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  scale_fill_grey() + theme_classic()

grid.arrange(p1, p2, ncol=2)
```



### 4. Change Legend

#### 4.1 Change Legend Position

The command `theme(legend.position="")` can be used to set the position of the legend. The legend also can be removed by this command.

```{r eval=F}
# Move the legend to top
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.position="top")

# Move the legend to bottom
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.position="bottom")

# Remove legend
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.position="none")
```

```{r echo=FALSE, fig.height=3.5, fig.width=10,message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Move the legend to top
p1 <- ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot()+
  theme(legend.position="top")

# Move the legend to bottom
p2 <- ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot()+
  theme(legend.position="bottom")

# Remove legend
p3 <- ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot()+
  theme(legend.position="none")

grid.arrange(p1, p2, p3, ncol=3)
```

__Move the legend manually__

The command `legend.justification` and `legend.position` can set the position of the legend manually. x,y is 0,0 (bottom left) to 1,1 (top right).


```{r eval=F}
# Put legend manually 
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.justification=c(0,0), legend.position=c(.5, .5))

# Put bottom-right corner of legend box in bottom-right corner of graph
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.justification=c(1,0), legend.position=c(1,0))
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

p1 <-ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.justification=c(0,0), legend.position=c(.5, .5))

# Put bottom-right corner of legend box in bottom-right corner of graph
p2 <- ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.justification=c(1,0), legend.position=c(1,0))

grid.arrange(p1, p2, ncol=2)
```



#### 4.2 Modifying the legend box

```{r eval=F}
# Remove the background of the ledgend box
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.background = element_rect())

# Change part of the color inside the legend
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.background = element_rect(fill="gray90", size=.5, linetype="dotted"))
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# Remove the background of the ledgend box
p1 <-ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.background = element_rect())

# Change part of the color inside the legend
p2 <-ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  theme(legend.background = element_rect(fill="gray90", size=.5, linetype="dotted"))

grid.arrange(p1, p2, ncol=2)
```


#### 4.3 Modifying the text of legend titles and labels

The `scale_fill_manual` command can change the color, legend titles and labels

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  # Using a manual scale instead 
  scale_fill_manual(values=c("#999999", "#E69F00"), 
                    name="Internet use \nPercentage",
                    breaks=c("OJ", "VC"),
                    labels=c("orange juice", "vitamin C"))
```

#### 4.4 Modifying legend title and labels

The following command change the text style of legend.

+ `legend.title = element_text()` : legend title.
+ `legend.text = element_text()` : legend categories.

The detial of __Formatting choices__ will be explain in __5.1__.

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
    # Using a manual scale instead 
  scale_fill_manual(values=c("#999999", "#E69F00"), 
                    name="Internet use \nPercentage",
                    breaks=c("OJ", "VC"),
                    labels=c("orange juice", "vitamin C"))+
  theme(legend.title = element_text(colour="blue", size=16, face="bold"),
        legend.text = element_text(colour="red", size = 14, face = "bold"))
```

### 5.Set Layout

#### 5.1 Change the Title and Axis Labels

There are two ways to change the Title and Axis Labels

+ `ylab`, `xlab` and `ggtitle` 
+ `labs(y='  ', x='  ', title='  ')`

```{r eval=F}
# first method to change the Title and Axis Labels
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point() +
  ylab('Gross Domestic Product') +
  xlab('Percentage of Gross Domestic \n Product Spent on Education') +
  ggtitle('Gross Domestic Product and Education')

# second method to change the Title and Axis Labels
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point() +
  labs(y='Domestic Product',
       x='Percentage of Product Spent on Education',
       title= 'Domestic Product and Education')
```

```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE, paged.print=FALSE}
require(gridExtra)

# first method to change the Title and Axis Labels
p1 <- ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point() +
  ylab('Gross Domestic Product') +
  xlab('Percentage of Gross Domestic \n Product Spent on Education') +
  ggtitle('Gross Domestic Product and Education')

# second method to change the Title and Axis Labels
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point() +
  labs(y='Domestic Product',
       x='Percentage of Product Spent on Education',
       title= 'Domestic Product and Education')

grid.arrange(p1, p2, ncol=2)
```


#### 5.2 Altering the text style of axis or title

The following command change the text style of axis and title

+ `plot.title = element_text()` : plot title.
+ `axis.title = element_text()` : axis title. 
+ `axis.text = element_text()` : appearance of axis values/numbers.

__Formatting choices__ The font, colour, size and emphasis of any of these labels can be altered by arguments within element_text(your format).

+ family: the font style. Examples of fonts include: “Palatino”, “Helvetica”, “Courier”, “Times”. For example, family = "Palatino"
+ face: the type of emphasis, with options including bold, italic and “bold.italic”. For example, face = "bold.italic"
+ colour: the colour can be changed to any of the colours listed here. For example, colour = "steelblue2".
+ size: the size of the text. For example, size = (3).


```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
# first method to change the Title and Axis Labels
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point() +
  ylab('Gross Domestic Product') +
  xlab('Percentage of Gross Domestic \n Product Spent on Education') +
  ggtitle('Gross Domestic Product and Education')+
  theme(axis.text=element_text(colour="blue", size=10), # change x and y label words
        axis.title=element_text(colour="orange", size=12, face="bold"),            # change x-axis and y-axis title words
        plot.title=element_text(colour="red", size=15, face="bold"))      # change title words size

```

#### 5.3 Removing a label

The command `element_blank()`, the following example remove the title and axis. This command also can remove the tiltle of legend. 

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
# first method to change the Title and Axis Labels
ggplot(data = CIACountries, aes(y=gdp, x = educ)) + geom_point() +
  ylab('Gross Domestic Product') +
  xlab('Percentage of Gross Domestic \n Product Spent on Education') +
  ggtitle('Gross Domestic Product and Education')+
  theme(axis.text= element_blank(),
        axis.title= element_blank(),
        plot.title= element_blank())    

```


#### 5.4 Scale

The `coord_trans()` function can put data on a logarithmic scale. A logarithmic scale would make the plot much more readable.

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data = CIACountries, aes(y=gdp, x = educ)) +
  geom_point() +
  guides(color=guide_legend(title="Internet Use"),
         size=guide_legend(title='Roadway')) +
  coord_trans(y = "log10")
```

#### 5.5 Facets

There alsp exist functions that allows the user to split one plot into multiple plots based on a factor included in the dataset.

+ `facet_wrap()` arranges a one-dimensional sequence of panels to allow them to cleanly fit on one page.
+ `facet_grid()` allows you to form a matrix of rows and columns of panels.

> `facet_wrap()` allows user to organise the panels as following:

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len)) +
  geom_boxplot(aes(fill = supp)) +
  facet_wrap(facets = vars(dose))
```

```{r fig.height=4, fig.width=5, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len)) +
  geom_boxplot(aes(fill = supp)) +
  facet_wrap(facets = vars(dose), ncol=2)
```


> `facet_grid()` allows user to organise the panels only by rows (or only by columns):

Split in horizontal direction

```{r fig.height=3.5, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  facet_grid(cols = vars(supp))
```

Split in vertical direction

```{r fig.height=3.5, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  facet_grid(rows = vars(supp))
```

Split in both vertical and horizontal direction

```{r fig.height=3.5, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Facet by two variables: dose and supp.
# Rows are dose and columns are supp
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  facet_grid(rows = vars(dose), cols = vars(supp))
```

### 6.Multiple plot

#### 6.1 Multiple plot 

Although facet can splitting one plot into multiple plots, sometimes users may want to produce a single figure that contains multiple plots using different variables or even different data frames. The __ggpubr__ package allows us to combine separate ggplots into a single figure using `ggarrange()`:

```{r  eval=F}
install.packages("ggpubr")
```

You then need to activate by executing the command:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggpubr)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
## 1. Create a box plot (p1)
p1 <- ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = supp), position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))

# 2. Create a scatter plot (p2)
p2 <- ggplot(data = CIACountries, aes(y=gdp, x = educ))  + geom_point(aes(color = net_users, size = roadways)) +
  guides(color=guide_legend(title="Internet Use"), size=guide_legend(title='Roadway'))


# 3. Create a box plot (p3)
p3 <- ggplot(ToothGrowth, aes(x = dose, y = len, fill=supp)) + geom_boxplot() +
  scale_fill_grey() + theme_classic() 
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
figure <- ggarrange(p1, p2, p3,
                    labels = c("A", "B", "C"), # label the plots 
                    ncol = 2, nrow = 2)
figure
```

#### 6.1 Multiple plot with layout

The nested `ggarrange()` functions can change column/row span of plots. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
ggarrange(
  p3,                # First row with box plot
  # Second row with box and scatter plots
  ggarrange(p1, p2, ncol = 2, labels = c("B", "C")), 
  nrow = 2, 
  labels = "A"       # Label of the line plot
  ) 
```


